---
title: "ST558 Project 1"
author: "Mu-Tien, Lee"
date: "2020,09.04"
output:
  github_document:
      toc: true
---

```{r setup, include=FALSE, linewidth=60}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

```

# Require Package
```{r require package}
#install.packages("qwraps2")
library(rmarkdown)
library(dplyr)
library(tidyverse)
library(knitr)
library(RSQLite)
library(bigrquery)
library(httr)
library(jsonlite)
library(ggplot2)
library(qwraps2)
```


# Data cleaning
## Building functions to reach statsAPI
```{r reaching statsAPI}
#build up base dataset for switch team.names and team.ID
text <- content(GET("https://statsapi.web.nhl.com/api/v1/teams?expand=team.roster"),"text")
base <- fromJSON(text, flatten = TRUE) 
base <- as.data.frame(base) %>% select(teams.franchise.teamName,teams.franchiseId,teams.teamName,teams.id)

#make a function to let user reach statsAPI
statsAPI <- function(x,teamID=NULL,season=NULL, ...){
  
#setting up base url for stats API
base_url <-"https://statsapi.web.nhl.com/api/v1/teams"
modifiers <- x

#construct the full path
if (x %in% "teamId"){
  full_url <- paste0(base_url, "?", modifiers, "=", teamID)
}
#convert team name into teamID
else if(is.character(teamID)){
  base <- base %>% filter(teams.teamName==teamID)
  teamID <- base[4]
}

if (x %in% c("expand=team.roster","expand=person.names","expand=team.schedule.next","expand=team.schedule.previous","expand=team.stats","stats=statsSingleSeasonPlayoffs")){
  
  if (is.null(teamID)){
    full_url <- paste0(base_url, "?", modifiers)}
  
  else {
    full_url <- paste0(base_url, "/", teamID, "?", modifiers)}
}  
  
else if (x %in% "expand=team.roster&season" ){
  if (is.null(season)){
    stop("seaon is missing")}
  
  else if(length(teamID)>1){
    stop("sorry, we can only show one team each time")}
  
  else if(is.null(teamID)){
    full_url <- paste0(base_url,"?", modifiers,"=", 20142015)}

  else {
    full_url <- paste0(base_url, "/", teamID, "?", modifiers,"=", season)}
}

# retrieve information in raw form
GET(full_url)
#transfer into JSON text form
text <- content(GET(full_url),"text")
#convert it to a list
mydata <- fromJSON(text, flatten = TRUE) 
mydata<- as.data.frame(mydata)
return(mydata)
}

#m<- statsAPI("expand=team.roster")
#n<- statsAPI("expand=person.names")
#o<- statsAPI("expand=team.schedule.next")
#p<- statsAPI("expand=team.schedule.previous")
#q<- statsAPI("expand=team.stats",12)
#r<- statsAPI("expand=team.roster&season",teamid=54, season = 19971998)
#s<- statsAPI("teamId","12,5,1")
#t<- statsAPI("stats=statsSingleSeasonPlayoffs")

```

## Building functions to reach recordAPI
```{r reaching recordAPI}
recordAPI <- function(x,franchiseID=NULL,...){
#setting up base url for stats API
base_url <-"https://records.nhl.com/site/api"
modifiers <- x

#convert team name into teamID
if(is.character(franchiseID)){
  base <- base %>% filter(teams.franchise.teamName==franchiseID)
  franchiseID <- base[2]
  franchiseID
}

#construct the full url path
if (is.null(franchiseID)){
    full_url <- paste0(base_url, "/", modifiers)
  }
else if (x %in% c("franchise","franchise-team-totals")) {
    stop("Sorry,table franchise and franchise-team-totals cannot indicate team")
  }
else {
  full_url <- paste0(base_url, "/", modifiers,"?cayenneExp=franchiseId=", franchiseID)
  } 

#retrieve information in raw form
GET(full_url)
#transfer into JSON text form
text <- content(GET(full_url),"text")
#convert it to a list
mydata <- fromJSON(text, flatten = TRUE) 
mydata<- as.data.frame(mydata)
return(mydata)
}

#a<-recordAPI("franchise")
#b<-recordAPI("franchise-team-totals")
#c<-recordAPI("franchise-season-records",11)
#d<-recordAPI("franchise-goalie-records",1)
#e<-recordAPI("franchise-skater-records",1)
```

Below is a function that can help you read whatever the endpoint you would like to get, please key in the following option in the function.
```{r describe the option of the function}
endpoints <- c("franchise","franchise-team-totals","franchise-season-records","franchise-goalie-records","franchise-skater-records","expand=team.roster","expand=person.names","expand=team.schedule.next","expand=team.schedule.previous","expand=team.stats","expand=team.roster&season","teamId","stats=statsSingleSeasonPlayoffs")

idOption <- c(rep("None",2),rep("franchiseID or teamName without location",3),rep("teamID or teamName without location",8))

seasonOption <- c(rep("No",10),"Yes", "No","No")

kable(cbind(endpoints,idOption,seasonOption), format = "html")
kable(base, format = "html",caption = "Id table for teams")
```


## Accessory for API endpoints
```{r accessory for endpoints}
endpoints <- function(x,...){
  record <- c("franchise","franchise-team-totals","franchise-season-records", "franchise-goalie-records", "franchise-skater-records")
  
  stats <- c("expand=team.roster","expand=person.names","expand=team.schedule.next","expand=team.schedule.previous","expand=team.stats","expand=team.roster&season","teamId","stats=statsSingleSeasonPlayoffs")
  
  if (x %in% record) recordAPI(x,...)
  else if (x %in% stats) statsAPI(x,...)
  else stop("Please enter the correct name of your enpoints")
}
#try<-endpoints("expand=person.names", teamID="Bruins")
```

# Data analysis

## Building data for later use
```{r build my data}
#get some wins/losses data for active teams
teamtotal<-endpoints("franchise-team-totals")%>% filter(is.na(data.lastSeasonId))

#get teams' division and conference from another table
division <- endpoints("expand=team.roster")%>% select(teams.id, teams.division.name, teams.division.nameShort, teams.conference.name) %>%rename(data.teamId=teams.id)

#combine them as new dataset
new <- left_join(teamtotal,division, by="data.teamId") 

#creat new variables
#calculating some rate
new <- new %>% mutate(winrate=data.wins/data.gamesPlayed, winlossrate=data.wins/data.losses, overtimelossrate= data.overtimeLosses/data.losses, goalpergame=data.goalsFor/data.gamesPlayed, averagePenaltytime= data.penaltyMinutes/data.gamesPlayed)%>% mutate(lossathomerate=data.homeLosses/(data.homeLosses+data.homeWins),winonraodrate=data.roadWins/(data.roadWins+data.roadLosses))

#put the level on the age of te team
for (i in 1: nrow(new)){
  
  if (new$data.firstSeasonId[i]/10000<1943){
    new$age[i]<- "Senior"
  }
  else if (new$data.firstSeasonId[i]/10000<1968){
    new$age[i]<- "Junior"
  }
  else if (new$data.firstSeasonId[i]/10000<1993){
    new$age[i]<- "Sophomore"
  }
  else {
    new$age[i]<- "Freshman"
  }
}

```

## Numeric summarize
### A contingency table (Division*Fist year of play)
```{r contingency table}
#level the age
new$age <- as.factor(new$age)
levels(new$age)<-list("1917-1942"="Freshman","1942-1967"="Sophomore","1967-1992"="Junior","1992-2017"="Senior")

#show the table
datafortable <- new %>% filter(data.gameTypeId==2)
table1 <- table(datafortable$age, datafortable$teams.division.name) 
kable(table1,caption = " Division and Team established year information") 

```
In this table, we can notice that Atlantic has younger team distribution when Centrl and Pacific have more older team

###Numeric summary table
```{r summary}
#make a summary table among all teams
calculate <- function(x,...){
  data <- new %>% filter(data.gameTypeId == x)%>% select(winrate,winlossrate,overtimelossrate,goalpergame)
  if (x==2) type <- "regular season" else type <- "play off season"

kable(apply(data,2, summary),format = "html", digit = 4,caption = paste0("Summary among all active teams during ", type))
}
calculate(2)
calculate(3)

```

```{r categorical summary}
#make a summary with grouping data by division
catesummary <- function(x,...){
  datacatesumm <- new %>% filter(data.gameTypeId == 2)%>% filter(teams.division.name == x) %>% select(data.goalsAgainst,data.goalsFor, data.points, data.pointPctg, data.wins)
    kable(apply(datacatesumm,2, summary), format="html", caption = paste("Summary of division", x), digit = 4)

}
catesummary("Metropolitan")
catesummary("Atlantic")
catesummary("Central")
catesummary("Pacific")
```

##Plots
### a bar plot tells conference and division
```{r bar plot}
#draw a bar plot to show the count in each division and conference
type2 <- new %>% filter(data.gameTypeId==2)
bar1 <- ggplot(data=type2,aes(x=teams.conference.name))
bar1+ geom_bar(aes(fill=teams.division.name), position = "dodge")+labs(title = "Barplot for teams' conference and division")+xlab("conference")
```


### a histogram of penalty times
```{r histogram}
#seperate data by regular season and play off season since the differnce are obvious
type3 <- new %>% filter(data.gameTypeId==3)

#make two histogram to show the distribution of penalty time
histo1 <- ggplot(data=type2,aes(x=averagePenaltytime))
histo1+geom_histogram(binwidth = 1, aes(fill=teams.division.name))+labs(title = "Histogram for penalty time during relugar season")+xlab("penalty time in minutes")

histo2 <- ggplot(data=type3,aes(x=averagePenaltytime))
histo2+geom_histogram(binwidth = 1, aes(fill=age))+labs(title = "Histogram for penalty time during play off season")+xlab("penalty time in minutes")
```

# data analysis between 4 selected team
I choose NY islanders, Tampa Bay lightening, Vegas Golden Knights, Dallas stars, which are in the Conference Finals in 2019-20 NHL season

```{r}
#filtering old dataset to only selected groups
fourteams <- filter(new, data.teamId %in% c(2, 14, 25, 54)) 
fourteams <- fourteams 

#making some new dataset on selected
#dataset of season record
NY <-endpoints("franchise-season-records",22)
Tampa_bay <-endpoints("franchise-season-records",31)
Vegas <-endpoints("franchise-season-records",38)
Dallas <-endpoints("franchise-season-records",15)
seasonrecord <- rbind.data.frame(NY, Tampa_bay, Vegas, Dallas)

#detail about the goalier of selected team
NY <-endpoints("franchise-goalie-records",22)
Tampa_bay <-endpoints("franchise-goalie-records",31)
Vegas <-endpoints("franchise-goalie-records",38)
Dallas <-endpoints("franchise-goalie-records",15)
goalierecord <- rbind.data.frame(NY, Tampa_bay, Vegas, Dallas)

#detail about the skater of selected team
NY <-endpoints("franchise-skater-records",22)
Tampa_bay <-endpoints("franchise-skater-records",31)
Vegas <-endpoints("franchise-skater-records",38)
Dallas <-endpoints("franchise-skater-records",15)
skaterrecord <- rbind.data.frame(NY, Tampa_bay, Vegas, Dallas)

#make one new variables for goalier
goalierecord <- goalierecord %>% mutate(winRate=data.wins/data.gamesPlayed)
```

##Numerical summary

```{r}
#rearrange the dataset
fourteams <- fourteams %>% select(data.teamName,data.gamesPlayed,winrate,winonraodrate,lossathomerate,data.gameTypeId)%>% arrange( data.gameTypeId)

#print it out
kable(fourteams[1:4,1:5], caption = "data summary of selected team during regular season")
kable(fourteams[5:8,1:5], caption = "data summary of selected team during play off season")

```


```{r}
#showing skater position distribution of selected between non-active player and avctive player
table2 <- table(skaterrecord$data.positionCode, skaterrecord$data.franchiseName, skaterrecord$data.activePlayer) 
kable(table2[ , ,1],format = "html",caption = " teams and skatter position for non-active player")
kable(table2[ , ,2],format = "html",caption = " teams and skatter position for active player")
```

##Plots
### scatter and boxplot 
```{r boxplot for activeskater}
#choosing player that have already played over 15 games
actskaterrecord <- skaterrecord %>%  filter(data.gamesPlayed > 15)

#make a boxplot for skater to see the summary of most point in one season
boxplot1 <- ggplot(data=actskaterrecord,aes(x=data.franchiseName, y=data.mostPointsOneSeason))
boxplot1+geom_boxplot()+labs(title="Boxplot of most point in one season for skater who palyed more than 15 games")

```

```{r boxplot for all goalier}
#showing the win rate boxplot of the goalier of selected team
boxplot2 <- ggplot(data=goalierecord,aes(x=data.franchiseName, y=winRate))
boxplot2+geom_boxplot()+geom_jitter(aes(color=data.franchiseName))+labs(title="Boxplot of win rate for goalier")
```

#render("Project1.Rmd", output_format ="github_document", output_file=I("README.md"))